---
title: 'STAT 443: Lab 7'
author: 'Saksham Sudershan (Student #31339427)'
date: "09/03/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tseries)
```

\section*{Question 1}

```{r, echo = TRUE}
data <- read.csv("TempPG.csv")
ts <- ts(data$Annual)
model1 <- arima(ts, order = c(1,0,0), include.mean = T)
model1
```
The fitted model would be:

$$ X_t - (-1.9591) = 0.5843 ( X_{t-1} +1.9591) + Z_t \; \; \; \; \; \; \; \; \;  where\ Z_t \sim \mathbb{N}\,(0,1.265) $$

\section*{Question 2}

```{r, echo = T}
acf(ts)
```
For an AR(1), we would expect the ACF to tail off exponentially or damped sinusoidally. However, here we see spikes at lags $h = 5, 10 \dots$ so this ACF graph differs from one of an AR(1) process.

\section*{Question 3}

```{r, echo = T}
ts_differenced <- diff(ts, lag = 1, differences = 1)
acf(ts_differenced, lag.max = 30)
pacf(ts_differenced, lag.max = 30)
```
The ACF of the differenced series cuts of at lag $h=1$ and the PACF tails off in no particular pattern, which means that we could fit an MA(1) model to this differenced series.

\section*{Question 4}

```{r, echo = T}
model2 <- arima(ts, order = c(0,1,1))
model2
```
The fitted ARIMA model is given by:

$$ X_t-X_{t-1} = Z_t -0.7504Z_{t-1} \; \; \; \; \; \; \; \; \;  where\ Z_t \sim \mathbb{N}\,(0,1.143) $$

\section*{Question 5}

```{r, echo = T}
tsdiag(model2)
```
The model appears to fit well. There seems to be very little correlation between the residuals, which is what we would want from our model. The p-values for the Ljung-Box statistic also are high for lags before $h=7$, which suggests that the fits reasonably well. 

\section*{Question 6}
```{r, echo = T}
model1$aic
model2$aic
```
As we can see here, the AIC value for the ARIMA(0,1,1) model is lower than that for the AR(1) model. Thus, we would select the ARIMA(0,1,1) model.