---
title: 'STAT 443: Lab 8'
author: 'Saksham Sudershan (Student #31339427)'
date: "14 March, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tseries)
library(stats)
```
\section*{Problem 1}
```{r, echo = T}
data <- read.delim("souvenir.txt", header = F)
ts <- ts(data$V1, start = c(1987,1), frequency = 12)
plot(ts, xlab = "Time", ylab = "Sales (in A$)", 
     main = "Monthly Sales from Jan 1987-Dec 1993")
acf(ts)
```
The series seems to have higher and higher peaks, which points to the existence of a multiplicative seasonal effect. The ACF clearly points out a period of 12 months or 1 year.

\section*{Problem 2}
```{r, echo = T}
#Extracting
prediction_ts <- window(ts, start = c(1987,1), end = c(1992, 12))
#Fitting prediction model using HoltWinters()
model_smooth <- HoltWinters(prediction_ts, seasonal = c("multiplicative"))
#Providing parameter values
model_smooth$alpha
model_smooth$beta
model_smooth$gamma
#Plotting
plot(model_smooth)
```
The Holt-Winter Smoothing parameters are $\alpha=$ `r model_smooth$alpha`, $\beta=$ `r model_smooth$beta` and $\gamma=$ `r model_smooth$gamma`.

\section*{Problem 3}
```{r, echo = T}
# Predicting for 1993
forecast_data <- data.frame(predict(model_smooth, 12, 
                                    level = 0.95, n.ahead = 12, prediction.interval = TRUE))
# Plotting
plot(forecast_data$fit, type = "l", xlab = "Time In Months", 
     ylab = "Sales (in A$)", main = "Predicted Sales for 1993")
for(i in 1:12){
  segments(x0 = i, y0 = forecast_data[i,2], x1 = i, y1= forecast_data[i,3],lty = 2, col = 3)
}
points(forecast_data$upr, pch = 4, col = 3)
points(forecast_data$lwr, pch = 4, col = 3)
legend("topleft", legend = c("Fitted Values", "95% Confidence Interval"), col = c(1, 3), 
       lty = c(1, 2), pch = c(2,4))
# Forecast for first 3 months
forecast_data[1:3,1]
```

\section*{Problem 4}
```{r, echo = T}
plot(ts[73:75], type = "l", xlab = "Time in Months" , ylab = "Sales (In A$)", 
     main = "Predicted Confidence Intervals and Observed values" )
for(i in 1:3){
  segments(x0 = i, y0 = forecast_data[i,2], x1 = i, y1= forecast_data[i,3],lty = 2, col = 3)
}
points(forecast_data$upr[1:3], pch = 4, col = 3)
points(forecast_data$lwr[1:3], pch = 4, col = 3)
legend("topleft", legend = c("Fitted Values", "95% Confidence Interval"), 
       col = c(1, 3), lty = c(1, 2), pch = c(NA,4))
```
The 95% CI interval for the Jan 1993 is `r forecast_data[1,3]` to `r forecast_data[1,2]` while the observed value is `r ts[73]`.

The 95% CI interval for the Feb 1993 is `r forecast_data[2,3]` to `r forecast_data[2,2]` while the observed value is `r ts[74]`.

The 95% CI interval for the March 1993 is `r forecast_data[3,3]` to `r forecast_data[3,2]` while the observed value is `r ts[75]`.


This can also be seen in the plot. Only the observed value of the first month lies within the 95% confidence interval.

\section*{Problem 5}

We could a consider a log transform to convert the model from a multiplicative seasonal effect to an additive seasonal effect. The results would be a time series which looks like this:

```{r, echo = T}
data2<- log(data)
ts1 <- ts(data2[,1])
plot(ts1, xlab = "Time", ylab = "Log of Sales (In A$)", main = "Transforming the 
     Series into an Additive Seasonal Effect Model")
```
