---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\pagenumbering{gobble}
\section*{Problem 5}

\subsection*{(a)}

For a Monte Carlo Approximation, we can simulate a sample of 20 8-sided die and counting the number of times we get \emph{i} rolls greater than or equal to 6 and \emph{j} rolls greater than or equal to 4. By counting these we can estimate the probability.

```{r eval=FALSE, include=T}
# Run number of simulations to estimate probability
  number_sim <- 10000
  
  # Add conditions to satisfy calculating probability for events
  condition_1 = condition_2 <- c(0)
  
  # Simulating n= number_sim rolls
  for(k in 1:number_sim){
    sample_rolls <- sample(1:8, size = 20, replace = TRUE)
    
    # Counting the number of dice rolls greater than or equal to 6 in the sample
    i_check <- sum(sample_rolls[]>=6)
    # Counting the number of dice rolls greater than or equal to 4 in the sample
    j_check <- sum(sample_rolls[]>=4)
    
    # Counting the samples which satisfy conditions 1, that is there are at least i rolls greater than 6
    if(i_check >= i){
      condition_1 = condition_1 + 1
    }
    # Counting the samples which satisfy conditions 1, that is there are at least j rolls greater than 4
    if(j_check >= j){
      condition_2 = condition_2 + 1
    }
  }
  # Estimating conditional probability
  prob <- condition_1/condition_2
```

\subsection{(b) and (c)}

We can create a function with inputs i and j to output the probabilities for the given i and j. Then we can run a loop to create a matrix for the probabilities for each i= [12,14] and j=[8,12].
```{r, echo=T}
sim_dice <- function(i,j){
  number_sim <- 10000
  condition_1 = condition_2 <- c(0)
  for(k in 1:number_sim){
    sample_rolls <- sample(1:8, size = 20, replace = TRUE)
    if(sum(sample_rolls[]>=6) >= i){
      condition_1 = condition_1 + 1
    }
    if(sum(sample_rolls[]>=4) >= j){
      condition_2 = condition_2 + 1
    }
  }
  prob <- condition_1/condition_2
  
  return(prob)
}

output <- data.frame(row.names = c("j=8","j=9","j=10","j=11","j=12"))
for(m in 1:3){
  for(o in 1:5){
    output[o,m] <- sim_dice(m+11,o+7)
  }
}
colnames(output) <- c("i=12","i=13","i=14")

output
```
The above are the estimated probabilities for every combination of \emph{i}s and \emph{j}s.