---
title: 'STAT 443: Lab 9'
author: 'Saksham Sudershan (Student #31339427)'
date: "21 March, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section*{Problem 1}
```{r, echo = T}
# Reading data and creating datasets
data("LakeHuron")
LakeHuron.train <- window(LakeHuron, start = 1875, end = 1967)
LakeHuron.test <- window(LakeHuron, start = 1968, end = 1972)

# Plotting training data
plot(LakeHuron.train, ylab = "Lake Levels (In Feet)", 
     main = "Lake Huron Level 1875-1967")
acf(LakeHuron.train, lag.max = 100)
pacf(LakeHuron.train, lag.max = 50)
```
The ACF tails off. The PACF cuts of lag $h=2$. This resembles an AR(2) process as the ACF tails off and the PACF cuts of at lag $p$.

\section*{Problem 2}
```{r, echo = T}
# Fitting an AR(2) process
LakeHuron.model <- arima(LakeHuron.train, order = c(2,0,0))
LakeHuron.model
```
The fitted model is:
$$ X_t - \mu = \alpha_1 (X_{t-1}- \mu) + \alpha_2 (X_{t-2}- \mu) + Z_t $$
$$ \Rightarrow X_t =578.9888+ 1.0627 (X_{t-1}- 578.9888) - 0.2691 (X_{t-2}-578.9888) + Z_t \; \; \; \; \; \; \; \; where\ Z_t \sim N(0,0.4815)$$

\section*{Problem 3}
```{r, echo = T}
# ACF of residuals
model.res <- ts(residuals(LakeHuron.model))
acf(model.res, lag.max = 100)

# Model Diagnostics
tsdiag(LakeHuron.model)
```
The residuals seem to have no significant autocorrelations between the lags.

The standardized residuals also seem to not follow any pattern and appear to resemble white noise. The p-values for Ljung-Box statistic is higher than 0.6 for most lags, thus we can comfortably say that this model fits fairly well.

\section*{Problem 4}
```{r, echo = T}
# Finding 95% CI
predicted.val <- data.frame(predict(LakeHuron.model, n.ahead = 3,
                                    level = 0.95, prediction.interval = T))
for(i in 1:3){
  predicted.val[i, 3] <- predicted.val[i,1]-1.96*predicted.val[i,2] 
  predicted.val[i, 4] <- predicted.val[i,1]+1.96*predicted.val[i,2]
}
colnames(predicted.val) <- c("pred", "se", "lwr", "upr") 
predicted.val
```

The forecast value for 1968 is `r round(predicted.val[1,1],4)`, and the 95% CI is (`r round(predicted.val[1,3],4)`,`r round(predicted.val[1,4],4)`).

The forecast value for 1969 is `r round(predicted.val[2,1],4)`, and the 95% CI is (`r round(predicted.val[2,3],4)`,`r round(predicted.val[2,4],4)`).

The forecast value for 1970 is `r round(predicted.val[3,1],4)`, and the 95% CI is (`r round(predicted.val[3,3],4)`,`r round(predicted.val[3,4],4)`).

\section*{Problem 5}
```{r, echo = T}
# Making plot of CI, predicted and actual values 
plot(c(1968, 1969, 1970),LakeHuron.test[1:3], type = "b", lty = 1, 
     col = 1, ylim = c(576.7,581.2), ylab = "Lake Levels(In Feet)", 
     xlab = "Time", main = "Forecast v. Actual Data")
points(c(1968, 1969, 1970), predicted.val[1:3,3], pch = 4, col = 4)
segments(x0=1968 ,y0=predicted.val[1,3],x1=1968,y1=predicted.val[1,4], lty = 3, col = 4)
segments(x0=1969,y0=predicted.val[2,3],x1=1969,y1=predicted.val[2,4], lty = 3, col = 4)
segments(x0=1970,y0=predicted.val[3,3],x1=1970,y1=predicted.val[3,4], lty = 3, col = 4)
points(c(1968, 1969, 1970), predicted.val[1:3,4], pch = 4, col = 4)
lines(c(1968, 1969, 1970),predicted.val[1:3,1], type = "b",lty = 2, col = 2, pch = 4)
legend("topleft", legend = c("Forecasted Values", "Actual","Conf. Interval"), 
       lty = c(2,1, 3), col = c(2,1, 4), pch = c(1, NA, 4))
```

The graph above shows that the predicted values are very close to the actual lake levels. This actual values also fall within the 95% confidence interval produced using our forecast. Thus, the forecast seems to work well.