---
title: 'STAT 443: Lab 1'
author: 'Saksham Sudershan (Student #31339427)'
date: "17 January, 2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tseries)
```

### Question 1

(a) 

```{r, echo=TRUE}
# Reading data
data <- read.csv("LakeLevels.csv")

# Looking at data
head(data,10)

# List of vectors in dataset
names(data)

# Creating plot
plot(data$LakeLevel, ylab = "Lake Levels", main = "Daily Lake Levels")
```
The plot is disjoint, and the x-axis is indexed for the time variable. For time series data, time should be on the x-axis and the points should be joined.


(b)

```{r, echo=TRUE}
# Checking whether dataframe is a time series object
is.ts(data)

# Fitting data into time series object
x <- ts(data = data$LakeLevel, start = c(2007,1), frequency = 365)
```


(c)

```{r, echo=TRUE}
# Plotting time series
plot(x, xlab="Time", ylab= "Lake Levels", main = "Daily Lake Levels")
```

The points are joint and the x-axis shows time in years. 
It has no discernible trend but does have a seasonality effect; the lake levels seem to increase every summer and go back down towards the winter of each year.

### Question 2

(a) 

```{r, echo=TRUE}
# Generating independent observations from standard normal distribution
set.seed(2022)
y <- rnorm(200)

# Fitting into class ts 
z <- ts(y)
```


(b)

```{r, echo=TRUE}
# Plotting
plot(z)
```

Since the standard deviation of our normal distribution is 1, values outside the range of 2 or -2 would be 2 standard deviations away from the mean which is 0. Using the empirical rule, we can say that we would expect only 5% of values to be outside, which is 10 values.

```{r, echo=TRUE}
# Counting no. of observations outside the range +-2
sum(abs(y)>2)
```


(c)

```{r, echo=TRUE}
# Creating sample autocorrelation function
acf(z, lag.max = 100, plot = TRUE)
```
The autocorrelation function shows that there is little correlation between past and present values. The autocorrelation for h=0 is 1, and almost all of the following autocorrelations fall within the 95% confidence limits and tends to stay within $\frac{\pm 2}{\sqrt{n}}$ . We can say that the data is random.
